---
title: "Chapter 02. Introduction: Credibility, Models, and Parameters"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

## 2.1. Bayesian inference is reallocation of credibility across possibilities

The code for Figure 2.1 is formidable. However, the bulk (i.e., the two `geom_text()` blocks and the subsequent `geom_line()` block) are just for the annotation in the lower three panels.

```{r, message = F, warning = F, fig.width = 6, fig.height = 4}
library(tidyverse)

tibble(
  iteration = rep(1:3, each = 2*4),
  stage = rep(rep(c("Prior", "Posterior"), each = 4), 
              times = 3),
  Possibilities = rep(LETTERS[1:4], times = 6),
  Credibility = c(rep(.25, times = 4),
                  c(0, rep(1/3, times = 3)),
                  c(0, rep(1/3, times = 3)),
                  rep(c(0, .5), each = 2),
                  rep(c(0, .5), each = 2),
                  c(rep(0, times = 3)), 1)
  ) %>%
  
  mutate(stage = factor(stage, levels = c("Prior", "Posterior"))) %>%
  
  # the plot
  ggplot(aes(x = Possibilities, 
             ymin = 0, ymax = Credibility)) +
  geom_linerange(size = 10, color = "grey30") +
  geom_text(data = tibble(
    Possibilities = "B",
    Credibility = .8,
    label = c("A is", "B is", "C is"),
    iteration = 1:3,
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
  ), aes(y = Credibility,
         label = label)
  ) +
  geom_text(data = tibble(
    Possibilities = "B",
    Credibility = .7,
    label = "impossible",
    iteration = 1:3,
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
  ), aes(y = Credibility,
         label = label)
  ) +
  geom_line(data = tibble(
    Possibilities = rep(LETTERS[1:3], each = 2),
    Credibility = rep(c(.63, 0), times = 3),
    iteration = rep(1:3, each = 2),
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
    ), aes(y = Credibility),
    arrow = arrow(length = unit(0.30,"cm"), 
                  ends = "last", type = "closed")) +
  ylab("Credibility") +
  facet_grid(stage ~ iteration) +
  theme(panel.grid = element_blank(),
        strip.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

Same deal for Figure 2.2.

```{r, fig.width = 2.25, fig.height = 4}
tibble(
  stage = rep(c("Prior", "Posterior"), each = 4),
  Possibilities = rep(LETTERS[1:4], times = 2),
  Credibility = c(rep(.25, times = 4),
                  c(rep(0, times = 3)), 1)
  ) %>%
  
  mutate(stage = factor(stage, levels = c("Prior", "Posterior"))) %>%
  
  ggplot(aes(x = Possibilities, 
             ymin = 0, ymax = Credibility)) +
  geom_linerange(size = 10, color = "grey30") +
  geom_text(data = tibble(
    Possibilities = "B",
    Credibility = .8,
    label = "D is",
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
  ), aes(y = Credibility,
         label = label)
  ) +
  geom_text(data = tibble(
    Possibilities = "B",
    Credibility = .7,
    label = "responsible",
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
  ), aes(y = Credibility,
         label = label)
  ) +
  geom_line(data = tibble(
    Possibilities = LETTERS[c(4, 4)],
    Credibility = c(.25, .98),
    stage = factor("Posterior", levels = c("Prior", "Posterior"))
    ), aes(y = Credibility),
    arrow = arrow(length = unit(0.30,"cm"), 
                  ends = "last", type = "closed"),
    color = "grey92") +
  ylab("Credibility") +
  facet_wrap(~stage, ncol = 1) +
  theme(panel.grid = element_blank(),
        axis.ticks.x = element_blank())
```

### 2.1.1. Data are noisy and inferences are probabilistic.

It's easier to break Figure 2.3 up into the top and bottom portions. Here's the top.

```{r, fig.width = 2.25, fig.height = 2.5}
d <-
  tibble(x = seq(from = -2, to = 6, by = .01))

mu_1 <- 1
mu_2 <- 2
mu_3 <- 3
mu_4 <- 4

ggplot(data = d, aes(x = x)) +
  # Note the use of mu_i in the data and aes() statements
  geom_ribbon(data = d %>% filter(x > mu_1 - .2 & x < mu_1 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_1, sd = 1.2)*.75),
              fill = "grey40") +
  geom_ribbon(data = d %>% filter(x > mu_2 - .2 & x < mu_2 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_2, sd = 1.2)*.75),
              fill = "grey40") +
  geom_ribbon(data = d %>% filter(x > mu_3 - .2 & x < mu_3 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_3, sd = 1.2)*.75),
              fill = "grey40") +
  geom_ribbon(data = d %>% filter(x > mu_4 - .2 & x < mu_4 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_4, sd = 1.2)*.75),
              fill = "grey40") +
  geom_line(aes(y = dnorm(x, mean = mu_1, sd = 1.2)*.75)) +  # Note the use of mu_i in the dnorm() statements
  geom_line(aes(y = dnorm(x, mean = mu_2, sd = 1.2)*.75)) +
  geom_line(aes(y = dnorm(x, mean = mu_3, sd = 1.2)*.75)) +
  geom_line(aes(y = dnorm(x, mean = mu_4, sd = 1.2)*.75)) +
  scale_x_continuous(breaks = 1:4) +
  coord_cartesian(xlim = 0:5,
                  ylim = 0:1) +
  labs(title = "Prior",
       x = "Possibilities", 
       y = "Credibility") +
  theme(panel.grid = element_blank(),
        axis.ticks.x = element_blank())
```

Now we're ready for the bottom.

```{r, fig.width = 2.25, fig.height = 2.5}
ggplot(data = d, aes(x = x)) +
  # Note the use of mu_i in the data and aes() statements
  geom_ribbon(data = d %>% filter(x > mu_1 - .2 & x < mu_1 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_1, sd = 1.2)*1/3),
              fill = "grey67") +
  geom_ribbon(data = d %>% filter(x > mu_2 - .2 & x < mu_2 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_2, sd = 1.2)*1.75),
              fill = "grey67") +
  geom_ribbon(data = d %>% filter(x > mu_3 - .2 & x < mu_3 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_3, sd = 1.2)*.925),
              fill = "grey67") +
  geom_ribbon(data = d %>% filter(x > mu_4 - .2 & x < mu_4 + .2),
              aes(ymin = 0, ymax = dnorm(x, mean = mu_4, sd = 1.2)*.05),
              fill = "grey67") +
  geom_line(aes(y = dnorm(x, mean = mu_1, sd = 1.2)*1/3)) +  # Note the use of mu_i in the dnorm() statements
  geom_line(aes(y = dnorm(x, mean = mu_2, sd = 1.2)*1.75)) +
  geom_line(aes(y = dnorm(x, mean = mu_3, sd = 1.2)*.925)) +
  geom_line(aes(y = dnorm(x, mean = mu_4, sd = 1.2)*.05)) +
  geom_point(data = tibble(x = c(1.75, 2.25, 2.75), y = 0),
             aes(x = x, y = y),
             size = 3, color = "grey33", alpha = 3/4) +
  scale_x_continuous(breaks = 1:4) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = .2)) +
  coord_cartesian(xlim = 0:5,
                  ylim = 0:1) +
  labs(title = "Posterior",
       x = "Possibilities", 
       y = "Credibility") +
  theme(panel.grid = element_blank(),
        axis.ticks.x = element_blank())
```

## 2.2. Possibilities are parameter values in descriptive models

Figure 2.4.a.

```{r, fig.width = 3, fig.height = 2.5}
set.seed(2.4)
d <- tibble(x = rnorm(2000, mean = 10, sd = 5))

ggplot(data = d, aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 1, fill = "grey67", 
                 color = "grey92", size = 1/10) +
  geom_line(data = tibble(x = seq(from = -6, to = 26, by = .01)),
            aes(x = x, y = dnorm(x, mean = 10, sd = 5)),
            color = "grey33") +
  coord_cartesian(xlim = -5:25) +
  labs(subtitle = "The candidate normal distribution\nhas a mean of 10 and SD of 5.",
       x = "Data Values", 
       y = "Data Probability") +
  theme(panel.grid = element_blank())
```

Figure 2.4.b.

```{r, fig.width = 3, fig.height = 2.5}
ggplot(data = d, aes(x = x)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 1, fill = "grey67",
                 color = "grey92", size = 1/8) +
  geom_line(data = tibble(x = seq(from = -6, to = 26, by = .01)),
            aes(x = x, y = dnorm(x, mean = 8, sd = 6)),
            color = "grey33", linetype = 2) +
  coord_cartesian(xlim = -5:25) +
  labs(subtitle = "The candidate normal distribution\nhas a mean of 8 and SD of 6.",
       x = "Data Values", 
       y = "Data Probability") +
  theme(panel.grid = element_blank())
```

## 2.3. The steps of Bayesian data analysis

In order to recreate Figure 2.5, we need to generate the data and fit the model. In his "HtWtDataDenerator" R script, Kruschke provided the code for a function that will generate height/weight data of the kind in his text. Here is the code:

```{r}
HtWtDataGenerator <- function(nSubj, rndsd = NULL, maleProb = 0.50) {
  # Random height, weight generator for males and females. Uses parameters from
  # Brainard, J. & Burmaster, D. E. (1992). Bivariate distributions for height and
  # weight of men and women in the United States. Risk Analysis, 12(2), 267-275.
  # Kruschke, J. K. (2011). Doing Bayesian data analysis:
  # A Tutorial with R and BUGS. Academic Press / Elsevier.
  # Kruschke, J. K. (2014). Doing Bayesian data analysis, 2nd Edition:
  # A Tutorial with R, JAGS and Stan. Academic Press / Elsevier.
  
  require(MASS)
  
  # Specify parameters of multivariate normal (MVN) distributions.
  # Men:
  HtMmu   <- 69.18
  HtMsd   <- 2.87
  lnWtMmu <- 5.14
  lnWtMsd <- 0.17
  Mrho    <- 0.42
  Mmean   <- c(HtMmu, lnWtMmu)
  Msigma  <- matrix(c(HtMsd^2, Mrho * HtMsd * lnWtMsd,
                      Mrho * HtMsd * lnWtMsd, lnWtMsd^2), nrow = 2)
  # Women cluster 1:
  HtFmu1   <- 63.11
  HtFsd1   <- 2.76
  lnWtFmu1 <- 5.06
  lnWtFsd1 <- 0.24
  Frho1    <- 0.41
  prop1    <- 0.46
  Fmean1   <- c(HtFmu1, lnWtFmu1)
  Fsigma1  <- matrix(c(HtFsd1^2, Frho1 * HtFsd1 * lnWtFsd1,
                       Frho1 * HtFsd1 * lnWtFsd1, lnWtFsd1^2), nrow = 2)
  # Women cluster 2:
  HtFmu2   <- 64.36
  HtFsd2   <- 2.49
  lnWtFmu2 <- 4.86
  lnWtFsd2 <- 0.14
  Frho2    <- 0.44
  prop2    <- 1 - prop1
  Fmean2   <- c(HtFmu2, lnWtFmu2)
  Fsigma2  <- matrix(c(HtFsd2^2, Frho2 * HtFsd2 * lnWtFsd2,
                       Frho2 * HtFsd2 * lnWtFsd2, lnWtFsd2^2), nrow = 2)
  
  # Randomly generate data values from those MVN distributions.
  if (!is.null(rndsd)) {set.seed(rndsd)}
  datamatrix <- matrix(0, nrow = nSubj, ncol = 3)
  colnames(datamatrix) <- c("male", "height", "weight")
  maleval <- 1; femaleval <- 0 # arbitrary coding values
  for (i in 1:nSubj)  {
    # Flip coin to decide sex
    sex <- sample(c(maleval, femaleval), size = 1, replace = TRUE,
                  prob = c(maleProb, 1 - maleProb))
    if (sex == maleval) {datum = mvrnorm(n = 1, mu = Mmean, Sigma = Msigma)}
    if (sex == femaleval) {
      Fclust = sample(c(1, 2), size = 1, replace = TRUE, prob = c(prop1, prop2))
      if (Fclust == 1) {datum = mvrnorm(n = 1, mu = Fmean1, Sigma = Fsigma1)}
      if (Fclust == 2) {datum = mvrnorm(n = 1, mu = Fmean2, Sigma = Fsigma2)}
    }
    datamatrix[i, ] = c(sex, round(c(datum[1], exp(datum[2])), 1))
  }
  
  return(datamatrix)
} # end function
```

Now we have the `HtWtDataGenerator()` function, all we need to do is determine how many values are generated and how probable we want the values to be based on those from men. These are controlled by the `nSubj` and `maleProb` parameters.

```{r, message = F, warning = F}
set.seed(57) # This makes the data generation reproducible

d_57 <-
  HtWtDataGenerator(nSubj = 57, maleProb = .5) %>%
  as_tibble()

d_57 %>%
  head()
```

Here's the model. We fit it with HMC via the [brms package](https://cran.r-project.org/web/packages/brms/index.html). The traditional use of [diffuse and noninformative priors is discouraged with HMC, as is the uniform distribution for sigma](https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations). Instead, we'll use weakly-regularizing priors for the intercept an slope and a half Cauchy with a fairly large scale parameter for $\sigma$.

```{r, message = F, warning = F}
library(brms)

k2.0 <- 
  brm(data = d_57, family = gaussian,
      weight ~ 1 + height,
      prior = c(set_prior("normal(0, 100)", class = "Intercept"),
                set_prior("normal(0, 100)", class = "b"),
                set_prior("cauchy(0, 10)", class = "sigma")),
      chains = 4, iter = 2000, warmup = 1000, cores = 4)

plot(k2.0)
print(k2.0)
```

We'll walk through all of this in greater detail starting in Chapter 8. But for now, here's Figure 2.5.a.

```{r, fig.width = 3.25, fig.height = 3}
post <- posterior_samples(k2.0)

n_lines <- 150

ggplot(data =  d_57, 
       aes(x = height, y = weight)) +
  geom_abline(intercept = post[1:n_lines, 1], 
              slope = post[1:n_lines, 2],
              color = "grey50", size = 1/4, alpha = .3) +
  geom_point(alpha = 2/3) +
  coord_cartesian(xlim = 55:80,
                  ylim = 50:250) +
  # the eval(substitute(paste())) trick came from: https://www.r-bloggers.com/value-of-an-r-object-in-an-expression/
  labs(subtitle = eval(substitute(paste("Data with", 
                                        n_lines, 
                                        "cridible regression lines"))),
       x = "Height in inches",
       y = "Weight in pounds") +
  theme(panel.grid = element_blank())
```

Figure 2.5.b. will take a little work. First, we'll use a custom function to compute the posterior mode and the `hdi()` function from the [HDInterval package](https://cran.r-project.org/web/packages/HDInterval/index.html) to compute the HPD intervals.
 
```{r}
# estimate_mode() comes from: https://stackoverflow.com/questions/2547402/is-there-a-built-in-function-for-finding-the-mode
estimate_mode <- function(x) {
  d <- density(x)
  d$x[which.max(d$y)]
  }

estimate_mode(post$b_height)

library(HDInterval)
hdi(post$b_height, credMass = .95) %>% as.vector
```

Figure 2.5.b.

```{r, fig.width = 3, fig.height = 3}
ggplot(data =  post, aes(x = b_height)) +
  geom_histogram(color = "grey92", fill = "grey50",
                 binwidth = .2, size = .2) +
  scale_y_continuous(NULL, breaks = NULL) +
  geom_point(data = tibble(b_height = estimate_mode(post$b_height)),
             aes(x = b_height, y = 0),
             size = 3) +
  geom_line(data = tibble(b_height = hdi(post$b_height, credMass = .95) %>% as.vector),
            aes(x = b_height, y = 0),
             size = 1.5) +
  coord_cartesian(xlim = 0:8) +
  labs(title = "The posterior distribution",
       subtitle = "The mode and 95% HPD intervals are depicted\nin the dot and horizontal line at the bottom.",
       x = expression(paste(beta[1], " (slope)"))) +
  theme(panel.grid = element_blank())
```

Figure 2.6.

```{r, fig.width = 4, fig.height = 4}
nd <- tibble(height = seq(from = 53, to = 81, length.out = 20))

pred_k2.0 <- 
  predict(k2.0, newdata = nd) %>%
  as_tibble() %>%
  bind_cols(nd)

ggplot(data =  pred_k2.0, aes(x = height)) +
  geom_pointrange(aes(y = Estimate, 
                      ymin = Q2.5, 
                      ymax = Q97.5),
                  color = "grey67",
                  shape = 20) +
  geom_point(data =  d_57, aes(x = height, y = weight),
             alpha = 2/3) +
  labs(subtitle = "Data with the percentile-based 95% intervals and the\nmeans of the posterior predictions",
       x = "Height in inches",
       y = "Weight in inches") +
  theme(panel.grid = element_blank())
```

The posterior predictions might be easier to depict with a ribbon and line, instead.

```{r, fig.width = 4, fig.height = 4}
nd <- tibble(height = seq(from = 53, to = 81, length.out = 30))

pred_k2.0 <- 
  predict(k2.0, newdata = nd) %>%
  as_tibble() %>%
  bind_cols(nd)

ggplot(data =  pred_k2.0, aes(x = height)) +
  geom_ribbon(aes(ymin = Q2.5, 
                  ymax = Q97.5),
                  fill = "grey75") +
  geom_line(aes(y = Estimate),
            color = "grey92") +
  geom_point(data =  d_57, aes(x = height, y = weight),
             alpha = 2/3) +
  labs(subtitle = "Data with the percentile-based 95% intervals and the\nmeans of the posterior predictions",
       x = "Height in inches",
       y = "Weight in inches") +
  theme(panel.grid = element_blank())
```


Note. The analyses in this document were done with:

* R          3.5.1
* RStudio    1.1.453
* rmarkdown  1.20
* tidyverse  1.2.1 
* MASS       7.3-50
* brms       2.3.4
* rstan      2.17.3
* HDInterval 0.2.0

## References
Kruschke, J. K. (2015). *Doing Bayesian data analysis, Second Edition: A tutorial with R, JAGS, and Stan.* Burlington, MA: Academic Press/Elsevier.

```{r, message = F, warning = F, echo = F}
# Here we'll remove our objects
rm(d, mu_1, mu_2, mu_3, mu_4, HtWtDataGenerator, d_57, k2.0, post, n_lines, estimate_mode, nd, pred_k2.0)
```